

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*  Load the Data
```{r setup, include=TRUE}
data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
```
*  Process/transform the data (if necessary) into a format suitable for your analysis
```{r setup, include=TRUE}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$interval <- factor(data$interval)
```
*  What is mean total number of steps taken per day?

* Missing values
```{r setup, include=TRUE}
indexes <- is.na(as.character(data$steps))
data_no_NA <- data[!indexes,]
head(data_no_NA)
```
*  Adding steps 
```{r setup, include=TRUE}
tstepsd <- aggregate(steps ~ date, data = data_no_NA, sum)
colnames(tstepsd) <- c("date", "steps")
```
* Histogram 
```{r setup, include=TRUE}
hist(as.numeric(tstepsd$steps), breaks = 10, col = "green", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")
```
*  Report the mean and median
```{r setup, include=TRUE}
mean(tstepsd$steps)
median(tstepsd$steps)
```


*  What is the average daily activity pattern?
*  Doing the average and ploting Calculating the average
```{r setup, include=TRUE}
stepspi <- aggregate(data_no_NA$steps, by=list(interval=data_no_NA$interval), FUN=mean)
colnames(stepspi) <- c("interval", "average_steps")
plot(as.integer(levels(stepspi$interval)), stepspi$average_steps, type="l",
     xlab = "Interval", ylab = "Average Number of Steps", main = "Average Daily Activity Pattern",  col ="green")
```
* Calculing the maximum number of steps
```{r setup, include=TRUE}
max_steps <- max(stepspi$average_steps)
max_steps
```
* Maximum number of steps with intervals
```{r setup, include=TRUE}
intervale_max_steps<-stepspi[which.max(stepspi$average_steps),]$interval
intervale_max_steps
```
*  Imputing missing values
*  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
```{r setup, include=TRUE}
sum(is.na(as.character(data$steps)))
sum(is.na(as.character(data$date)))
sum(is.na(as.character(data$interval)))
```
* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc

```{r setup, include=TRUE}
indexes <- which(is.na(as.character(data$steps)))
complete_data <- data
complete_data[indexes, ]$steps<-unlist(lapply(indexes, FUN=function(indexes){stepspi[data[indexes,]$interval==stepspi$interval,]$average_steps}))

```
*  Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r setup, include=TRUE}
summary(complete_data)
str(complete_data)
```
*  Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r setup, include=TRUE}
tstepsd_complete <- aggregate(steps ~ date, data = complete_data, sum)
colnames(tstepsd_complete) <- c("date", "steps")
hist(as.numeric(tstepsd_complete$steps), breaks = 20, col = "green", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")
```

```{r setup, include=TRUE}
mean(tstepsd_complete$steps)
median(tstepsd_complete$steps)
```
*  Are there differences in activity patterns between weekdays and weekends?
* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r setup, include=TRUE}
complete_data$day <- as.factor(weekdays(complete_data$date))
complete_data$is_weekday <- ifelse(!(complete_data$day %in% c("Saturday","Sunday")), TRUE, FALSE)
weekdays_data <- complete_data[complete_data$is_weekday,]
stepspi_weekdays <- aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval), FUN=mean)

weekends_data <- complete_data[!complete_data$is_weekday,]
stepspi_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)

colnames(stepspi_weekdays) <- c("interval", "average_steps")
colnames(stepspi_weekends) <- c("interval", "average_steps")

stepspi_weekdays$day <- "Weekday"
stepspi_weekends$day <- "Weekend"

week_data <- rbind(stepspi_weekends, stepspi_weekdays)

week_data$day <- as.factor(week_data$day)
```
*  Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r setup, include=TRUE}

library(lattice)
xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")

```

 


  

